\texttt{davos}: a Python package ``smuggler'' for constructing lightweight reproducible notebooks

### 2022-10-27 paper/main.tex
updates through line 149 (beginning of 2.2.3, davos config)

% ======== additional package not in SoftwareX template ========
\usepackage[colorlinks=true, urlcolor=blue, pdfborder={0 0 0}]{hyperref}
\usepackage{breakurl}
\usepackage[htt]{hyphenat}
%\usepackage{listings}
%\usepackage{xcolor}
%\usepackage{todonotes}

@@

\subsubsection{The onion comment}\label{subsec:onion}
For greater control over the behavior of \texttt{smuggle} statements, \texttt{davos} defines an additional construct called the ``onion comment.'' An onion comment is a special type of inline comment that may be placed on a line containing a \texttt{smuggle} statement to customize how \texttt{davos} searches for the smuggled package locally and, if necessary, downloads and installs it. Onion comments follow a simple format based on the ``type comment'' syntax introduced in PEP 484~\cite{vanREtal14}, and are designed to make managing packages with \texttt{davos} intuitive and familiar. To construct an onion comment, users provide the name of the installer program (e.g., \texttt{pip}) and the same arguments one would use to manually install the package as desired via the command line:
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/snippet1}
\end{center}
Onion comments are useful when smuggling a package whose distribution name (i.e., the name used when installing it) is different from its top-level module name (i.e., the name used when importing it):
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/snippet2}
\end{center}
However, the most powerful use of the onion comment is making \texttt{smuggle} statements \textit{version-sensitive}. If an onion comment includes a version specifier~\cite{CoghStuf13}, \texttt{davos} will ensure that the version of the package loaded into the notebook matches the specific version requested, or satisfies the given version constraints. If the smuggled package exists locally, \texttt{davos} will extract its version info from its metadata and compare it to the specifier provided. If the two are incompatible (or no local installation is found), \texttt{davos} will install and load a suitable version of the package instead:
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/snippet3}
\end{center}
Onion comments can similarly be used to smuggle specific VCS references (e.g., Git~\cite{TorvHama05}  branches, commits, tags, etc.):
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/snippet4}
\end{center}
\texttt{davos} processes onion comments internally before forwarding arguments to the installer program. In addition to preventing onion comments from being used as a vehicle for shell injection attacks, this allows \texttt{davos} take certain logical actions when particular arguments are passed. For example, the \texttt{-I}/\texttt{--ignore-installed}, \texttt{-U}/\texttt{--upgrade}, and \texttt{--force-reinstall} flags will all cause \texttt{davos} to skip searching for a smuggled package locally before installing a new copy:
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/snippet5}
\end{center}
Similarly, passing \texttt{--no-input} will temporarily enable \texttt{davos}'s non-interactive mode (see Section~\ref{subsec:onion}), and installing a smuggled package into \texttt{<dir>} with \texttt{--target <dir>} will cause \texttt{dir} to be prepended to the module search path (\texttt{sys.path}), if necessary, so the package can be imported

\subsubsection{The \texttt{davos} config}\label{subsec:config}
The \texttt{davos} config object provides a simple, high-level interface that allows users to view and set various options that affect \texttt{davos}'s behavior. After importing \texttt{davos}, the config instance (a singleton) for the current session is available as \texttt{davos.config}, and its various fields are accessible as attributes. The config object exposes a mixture of writable and read-only fields. Writable fields include:
\begin{itemize}
\item \texttt{.active}: Whether or not \texttt{davos} functionality (i.e., support for \texttt{smuggle} statements and onion comments) should be enabled for subsequent code. Defaults to \texttt{True} when \texttt{davos} is first imported. See Section~\ref{subsec:implementation} for additional info.
\item \texttt{.auto\_rerun}: Controls behavior if \texttt{davos} is used to \texttt{smuggle} a new version of a package that was previously imported and cannot be reloaded (i.e., it contains C-extensions that dynamically generate code). If \texttt{True} (default: \texttt{False}), \texttt{davos} will automatically restart the notebook kernel and rerun all code up to (and including) the current \texttt{smuggle} statement. Otherwise, \texttt{davos} will issue a warning, pause execution, and prompt the user with buttons to either restart \& rerun the notebook or continue running with the imported package version. (Note: not configurable in Google Colaboratory).
\item \texttt{.confirm\_install}: If \texttt{True} (default: \texttt{False}), \texttt{davos} will require user confirmation (\texttt{[y]es/[n]o} input) before installing a smuggled package.
\item \texttt{.noninteractive}: Setting to \texttt{True} (default: \texttt{False}) enables non-interactive mode, in which all user input and confirmation is disabled. Note that in non-interactive mode, the \texttt{confirm\_install} option is set to \texttt{False}, and if \texttt{auto\_rerun} is \texttt{False}, \texttt{davos} will throw an error if a smuggled package cannot be reloaded.
\item \texttt{.pip\_executable}: The path to the \texttt{pip} executable used to install smuggled packages. Default is programmatically determined from the Python environment and falls back to \texttt{sys.executable -m pip} if one can't be found.
\item \texttt{.suppress\_stdout}: If \texttt{True} (default: \texttt{False}), suppress all unnecessary output issued by both \texttt{davos} and the installer program. Useful when smuggling packages that need to install many dependencies and/or generate extensive output. If the installer program throws an error, both stdout and stderr will be shown with the traceback.
\end{itemize}

>>>>

% ======== additional package not in SoftwareX template ========
\usepackage[colorlinks=true, urlcolor=blue, pdfborder={0 0 0}]{hyperref}
\usepackage{breakurl}
\usepackage[htt]{hyphenat}
%\usepackage{listings}
\usepackage{xcolor}
%\usepackage{todonotes}

@@

\subsubsection{The onion comment}\label{subsec:onion}
For greater control over the behavior of \texttt{smuggle} statements, \texttt{davos} defines an additional construct called the ``onion comment.'' An onion comment is a special type of inline comment that may be placed on a line containing a \texttt{smuggle} statement to customize how \texttt{davos} searches for the smuggled package locally and, if necessary, downloads and installs it. Onion comments follow a simple format based on the ``type comment'' syntax introduced in PEP 484~\cite{vanREtal14}, and are designed to make managing packages with \texttt{davos} intuitive and familiar. To construct an onion comment, users provide the name of the installer program (e.g., \texttt{pip}) and the same arguments one would use to manually install the package as desired via the command line:
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/snippet1}
\end{center}
Occasionally, a package's distribution name (i.e., the name used when installing it) may differ from its top-level module name (i.e., the name used when importing it). In such cases, an onion comment can be used to ensure \texttt{davos} installs the proper distribution if the smuggled package can't be found locally:
%Onion comments are useful in such cases for ensuring \texttt{davos} installs the proper distribution if the smuggled package isn't installed locally:
%Onion comments are useful when smuggling a package whose distribution name (i.e., the name used when installing it) differs from its top-level module name (i.e., the name used when importing it). Onion comments are useful in such cases for 
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/snippet2}
\end{center}
However, the most powerful use of the onion comment is making \texttt{smuggle} statements \textit{version-sensitive}. If an onion comment includes a version specifier~\cite{CoghStuf13}, \texttt{davos} will ensure that the version of the package loaded into the notebook matches the specific version requested, or satisfies the given version constraints. If the smuggled package exists locally, \texttt{davos} will extract its version info from its metadata and compare it to the specifier provided. If the two are incompatible (or no local installation is found), \texttt{davos} will install and load a suitable version of the package instead:
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/snippet3}
\end{center}
Onion comments can also be used to smuggle specific VCS references (e.g., Git~\cite{TorvHama05}  branches, commits, tags, etc.):
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/snippet4}
\end{center}
\texttt{davos} processes onion comments internally before forwarding arguments to the installer program. In addition to preventing onion comments from being used as a vehicle for shell injection attacks, this allows \texttt{davos} to take certain logical actions when particular arguments are passed. For example, the \texttt{-I}/\newline\texttt{--ignore-installed}, \texttt{-U}/\texttt{--upgrade}, and \texttt{--force-reinstall} flags will all cause \texttt{davos} to skip searching for a smuggled package locally before installing a new copy:
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/snippet5}
\end{center}
Similarly, passing \texttt{--no-input} will temporarily enable \texttt{davos}'s non-interactive mode (see Section~\ref{subsec:onion}), and installing a smuggled package into \texttt{<dir>} with \texttt{--target <dir>} will cause \texttt{dir} to be prepended to the module search path (\texttt{sys.path}), if necessary, so the package can be imported

\subsubsection{The \texttt{davos} config}\label{subsec:config}
The \texttt{davos} config object provides a simple, high-level interface that allows users to view and set various options that affect \texttt{davos}'s behavior. After importing \texttt{davos}, the config instance (a singleton) for the current session is available as \texttt{davos.config}, and its various fields are accessible as attributes (see Fig.~\textcolor{red}{REF TO ILLUSTRATIVE EXAMPLE FIG} for example usage). The config object exposes a mixture of writable and read-only fields. Writable fields include:
\begin{itemize}
\item \texttt{.active}: Whether or not \texttt{davos} functionality (i.e., support for \texttt{smuggle} statements and onion comments) should be enabled for subsequent code. Defaults to \texttt{True} when \texttt{davos} is first imported. See Section~\ref{subsec:implementation} for additional info.
\item \texttt{.auto\_rerun}: Controls behavior if \texttt{davos} is used to \texttt{smuggle} a new version of a package that was previously imported and cannot be reloaded (i.e., it contains C-extensions that dynamically generate code). If \texttt{True} (default: \texttt{False}), \texttt{davos} will automatically restart the notebook kernel and rerun all code up to (and including) the current \texttt{smuggle} statement. Otherwise, \texttt{davos} will issue a warning, pause execution, and prompt the user with buttons to either restart \& rerun the notebook or continue running with the imported package version. (Note: not configurable in Google Colaboratory).
\item \texttt{.confirm\_install}: If \texttt{True} (default: \texttt{False}), \texttt{davos} will require user confirmation (\texttt{[y]es/[n]o} input) before installing a smuggled package.
\item \texttt{.noninteractive}: Setting to \texttt{True} (default: \texttt{False}) enables non-interactive mode, in which all user input and confirmation is disabled. Note that in non-interactive mode, the \texttt{confirm\_install} option is set to \texttt{False}, and if \texttt{auto\_rerun} is \texttt{False}, \texttt{davos} will throw an error if a smuggled package cannot be reloaded.
\item \texttt{.pip\_executable}: The path to the \texttt{pip} executable used to install smuggled packages. Default is programmatically determined from the Python environment and falls back to \texttt{sys.executable -m pip} if one can't be found.
\item \texttt{.suppress\_stdout}: If \texttt{True} (default: \texttt{False}), suppress all unnecessary output issued by both \texttt{davos} and the installer program. Useful when smuggling packages that need to install many dependencies and/or generate extensive output. If the installer program throws an error, both stdout and stderr will be shown with the traceback.
\end{itemize}
